Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming

### 2014-12-18 manuscripts/manuscript.Rmd
fix pandoc silent parsing error creating NA

\begin{equation}
E = K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I}_n \right)^{-1} \mathbf{Y}_{\textrm{obs}}
\end{equation}

@@

\begin{equation}
C = K(X_p, X_p) - K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, X_p)
\end{equation}

Here $\mathbb{I}_n$ is the $n$ by $n$ identity matrix (i.e. a matrix with ones down the diagonal and zeros elsewhere) and $K$ is the ‘covariance kernel.’  The covariance kernel controls how much influence one observation has on another.  In the present application we use the squared-exponential kernel which, when evaluated over a pair of vectors, say $\mathbf{x}$ and $\mathbf{y}$,  generates a covariance matrix whose $i,j$th element is given by

@@

### Traditional model-choice approaches can be positively misleading.

Our results illustrate that model-choice approaches can be absolutely
misleading – by providing support to models that cannot capture tipping point
dynamics because they have fewer parameters and the data are far from the tipping point.  That is, when the data 
come from around the stable steady state,
all the parametric models are approximately linear and approximately
identical.  Thus, it is intuitive that all model selection methods choose the simplest model.
In a complex world, the result is suboptimal. But in a world that might
contain tipping points, the result could be disastrous.

>>>>

\begin{equation}
E = K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbf{I}_n \right)^{-1} \mathbf{Y}_{\textrm{obs}}
\end{equation}

@@

\begin{equation}
C = K(X_p, X_p) - K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbf{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, X_p)
\end{equation}

Here $\mathbf{I}_n$ is the $n$ by $n$ identity matrix (i.e. a matrix with ones down the diagonal and zeros elsewhere) and $K$ is the ‘covariance kernel.’  The covariance kernel controls how much influence one observation has on another.  In the present application we use the squared-exponential kernel which, when evaluated over a pair of vectors, say $\mathbf{x}$ and $\mathbf{y}$,  generates a covariance matrix whose $i,j$th element is given by

@@

### Traditional model-choice approaches can be positively misleading

Our results illustrate that model-choice approaches can be absolutely
misleading -- by providing support to models that cannot capture tipping point
dynamics because they have fewer parameters and the data are far from the tipping point.  That is, when the data 
come from around the stable steady state,
all the parametric models are approximately linear and approximately
identical.  Thus, it is intuitive that all model selection methods choose the simplest model.
In a complex world, the result is suboptimal. But in a world that might
contain tipping points, the result could be disastrous.

### 2014-12-18 manuscripts/manuscript.tex
fix zenodo ref
pdf figures
fix pandoc silent parsing error creating NA

\subsubsection{Parametric models}\label{parametric-models}

@@

\subsubsection{Parametric models}\label{parametric-models}

@@

\begin{equation}
C = K(X_p, X_p) - K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, X_p)
\end{equation}

Here $\mathbb{I}_n$ is the $n$ by $n$ identity matrix (i.e.~a matrix
with ones down the diagonal and zeros elsewhere) and $K$ is the
`covariance kernel.' The covariance kernel controls how much influence
one observation has on another. In the present application we use the
squared-exponential kernel which, when evaluated over a pair of vectors,
say $\mathbf{x}$ and $\mathbf{y}$, generates a covariance matrix whose
$i,j$th element is given by

@@

For the prior on $\ell$, $\alpha = $ 10 and $\beta = $ 10. The prior on
$\sigma$, $\alpha = $ 5 and $\beta = $ 5.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2-1.jpeg}
\caption{The inferred Gaussian process compared to the true process and
maximum-likelihood estimated (MLE) process. We show the expected value
for the function $f$ under each model. Plots show replicates drawn from
the posterior distributions in order to convey uncertainty of the
estimates. Note the MLE is a point estimate of parameters and so
reflects no uncertainty in the distribution. The training data are also
shown as black points. The GP is conditioned on (0,0), shown as a
pseudo-data point.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_4-1.jpeg}
\caption{In the management context, GPDP outperforms approaches based on
parametric models. We show 100 replicate simulations of the stock
dynamics (Eqn 1) under the policies derived from each of the estimated
models, as well as the policy based on the exact underlying model.}
\end{figure}

@@

\subsubsection{Traditional model-choice approaches can be positively
misleading.}\label{traditional-model-choice-approaches-can-be-positively-misleading.}

@@

>>>>

@@

\subsubsection{Parametric Models}\label{parametric-models}

@@

\begin{equation}
C = K(X_p, X_p) - K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbf{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, X_p)
\end{equation}

Here $\mathbf{I}_n$ is the $n$ by $n$ identity matrix (i.e.~a matrix
with ones down the diagonal and zeros elsewhere) and $K$ is the
`covariance kernel.' The covariance kernel controls how much influence
one observation has on another. In the present application we use the
squared-exponential kernel which, when evaluated over a pair of vectors,
say $\mathbf{x}$ and $\mathbf{y}$, generates a covariance matrix whose
$i,j$th element is given by

@@

For the prior on $\ell$, $\alpha =$ 10 and $\beta =$ 10. The prior on
$\sigma$, $\alpha =$ 5 and $\beta =$ 5.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2-1.pdf}
\caption{The inferred Gaussian process compared to the true process and
maximum-likelihood estimated (MLE) process. We show the expected value
for the function $f$ under each model. Plots show replicates drawn from
the posterior distributions in order to convey uncertainty of the
estimates. Note the MLE is a point estimate of parameters and so
reflects no uncertainty in the distribution. The training data are also
shown as black points. The GP is conditioned on (0,0), shown as a
pseudo-data point.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_4-1.pdf}
\caption{In the management context, GPDP outperforms approaches based on
parametric models. We show 100 replicate simulations of the stock
dynamics (Eqn 1) under the policies derived from each of the estimated
models, as well as the policy based on the exact underlying model.}
\end{figure}

@@

\subsubsection{Traditional model-choice approaches can be positively
misleading}\label{traditional-model-choice-approaches-can-be-positively-misleading}

@@

Xie, Y. 2013. Dynamic Documents with R and knitr. Chapman; Hall/CRCBoca
Raton, Florida.
